Power to the Stimuli: Not the Effect

### 2023-08-03 manuscript_draft.Rmd
update the power

---
title             : "Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli"
shorttitle        : "Title"

@@

  Sample size planning for research studies often focuses on obtaining a significant result given a specified level of power, alpha, and proposed effect size. This planning generally requires prior knowledge of study design and a statistical analysis to calculate the proposed sample size. However, there may not be just one specific testable analysis from which to derive power [@silberzahn2018many] or even a hypothesis to test for the project (e.g., stimuli database creation). Newer power and sample size planning suggestions include Accuracy in Parameter Estimation [AIPE, @kelley2007, @maxwell2008] and simulation of proposed analyses [@chalmers2020]. These toolkits provide flexibility in traditional power analyses that focus on the if-this-then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size. In this tutorial, we explore how these latter two approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses. Specifically, the examples focus on studies that implement multiple items and suggest that sample sizes can be planned to measure those items adequately and accurately, regardless of statistical test. Results show that pilot data can be used to determine a sample size that represents well-measured data, and multiple code vignettes will be provided for researchers to adapt and apply to their own measures.

  The planning of the sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size. This planning requires prior knowledge of the study design and a statistical analysis to calculate the proposed sample size. However, there may not be one specific testable analysis from which to derive power (Silberzahn et al., 2018) or a hypothesis to test for the project (e.g., creation of a stimuli database). Modern power and sample size planning suggestions include accuracy in parameter estimation (AIPE; Maxwell et al., 2008) and simulation of proposed analyses (Chalmers & Adkins, 2020). These toolkits provide flexibility in traditional power analyses that focus on the if-this, then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size. In this tutorial, we explore how AIPE and simulation approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses. Specifically, the examples focus on studies that adopt multiple items and suggest that sample sizes can be planned to measure those items adequately and precisely, regardless of statistical test. We demonstrate that pilot data can be used to determine a sample size that represents well-measured data.  This tutorial also provides multiple code vignettes that researchers can adapt and apply to their own measures.

@@

Statistical power and power analyses are arguably one of the most important components to planning a research study [@cohen1990]. Yet, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts [@hardwicke2020, @doi:10.1177/1745691620979806]. The replication "crisis" and credibility revolution have shown that many published studies in psychology are underpowered [@opensciencecollaboration2015, @vazire2018]. Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020, @d.chambers2014, @nosek2014] and then should summarily include a power analysis to determine the sample size necessary to detect the expected effect. Given the combined issues of publish-or-perish and that most non-significant results do not turn in published manuscripts, one may expect that power analysis would be especially critical for early career researchers [@rosenthal1979, @simmons2011]. Potentially, it is uninformative to publish studies that are underpowered [@halpern2002], but it can be difficult to know if a bad power analysis is better than no power analysis. A recent review of power analyses found in psychology journal articles indicates that researchers did not provide enough information to understand their power analyses and often chose effect sizes that were inappropriately justified [@beribisky2019].

One potential solution to the power analysis problem is the plethora of tools made available for researchers to simply power. G\*Power is one of the most popular free power software options [@faul2007, @erdfelder1996] that provides a simple point and click graphical user interface for power. Web-based tools have also sprung up for overall and statistical test specific sample size planning including powerandsamplesize.com and <https://designingexperiments.com> [@anderson2017]. Coding based packages, such as *pwr* [@champely2017], *faux* [@debruine2021], and *SimDesign* [@chalmers2020] can be used to examine power and sample size planning using *R*, usually with simulation. Researchers have to be careful using any toolkit, as errors can occur with the over-reliance on software [@nuijten2016]. When computing sample size estimates it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result [@williamson2000, @batterham2005].

@@

While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques. Therefore, we provide XX example vignettes and code examples on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript. Examples include psycholinguistics, social psychology, COVID related data, and cognitive psychology.

>>>>

---
title             : "Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning with Multiple Stimuli"
shorttitle        : "AIPE SIMULATION"

@@

  The planning of the sample size for research studies often focuses on obtaining a significant result given a specified level of power, significance, and an anticipated effect size. This planning requires prior knowledge of the study design and a statistical analysis to calculate the proposed sample size. However, there may not be one specific testable analysis from which to derive power [@silberzahn2018many] or a hypothesis to test for the project (e.g., creation of a stimuli database). Modern power and sample size planning suggestions include accuracy in parameter estimation [AIPE, @kelley2007; @maxwell2008] and simulation of proposed analyses [@chalmers2020]. These toolkits provide flexibility in traditional power analyses that focus on the if-this, then-that approach, yet, both AIPE and simulation require either a specific parameter (e.g., mean, effect size, etc.) or statistical test for planning sample size. In this tutorial, we explore how AIPE and simulation approaches can be combined to accommodate studies that may not have a specific hypothesis test or wish to account for the potential of a multiverse of analyses. Specifically, the examples focus on studies that adopt multiple items and suggest that sample sizes can be planned to measure those items adequately and precisely, regardless of statistical test. We demonstrate that pilot data can be used to determine a sample size that represents well-measured data. This tutorial also provides multiple code vignettes that researchers can adapt and apply to their own measures.

@@

Statistical power and power analyses are arguably one of the most important components in planning a research study [@cohen1990]. However, if reviews of transparency and openness in research publications are any clue, the social sciences have failed to fully implement power analyses as part of their common efforts [@hardwicke2020; @doi:10.1177/1745691620979806]. The replication \"crisis\" and credibility revolution have shown that published studies in psychology are underpowered [@opensciencecollaboration2015; @vazire2018]. Pre-registration of a study involves outlining the study and hypotheses before data collection begins [@stewart2020; @d.chambers2014; @nosek2014], and details of a power analyses or other limitations on resources are often used to provide justification for the pre-registered sample quota. Given the combined issues of publish-or-perish and that most non-significant results do not result in published manuscripts, one may expect that power analysis would be especially critical for early career researchers [@rosenthal1979; @simmons2011]. At best, an underpowered study provides limited insight [@halpern2002], and it can be difficult to know if a poorly implemented power analysis is better than no power analysis.

A recent review of power analyses found - across psychology journal articles - researchers did not provide enough information to understand their power analyses and often chose effect sizes that were unjustified [@beribisky2019]. One solution to this power analysis problem is the plethora of tools made available for researchers to make power computations accessible to non-statisticians with the caveat that a solid education in power is necessary to use these tools. G\*Power is one of the most popular free power software options [@faul2007; @erdfelder1996] that provides a simple point and click graphical user interface for power calculations [however, see @brysbaert]. Web-based tools have also sprung up for overall and statistical test specific sample size planning including <https://powerandsamplesize.com> and <https://designingexperiments.com> [@anderson2017]. *R*-coding based packages, such as *pwr* [@champely2017], *faux* [@debruine2021], and *SimDesign* [@chalmers2020], can be used to examine power and sample size planning usually with simulation. Researchers must be careful using any toolkit, as errors can occur with the over-reliance on software [e.g., it should not be a substitute for critical thinking, @nuijten2016]. When computing sample size estimates, it is important to remember that these values are estimations, not exact calculations guaranteed to produce a specific result [@batterham2005]. For example, if any parameter estimated by the researcher was not found in the study (i.e., a smaller effect size than used for the power analysis), then the sample size estimate tied to that specific level of power may be incorrect.

STOPPED HERE

@@

While the examples in this manuscript are traditionally cognitive linguistics focused, any research using repeated items can benefit from newer sampling techniques. Therefore, we provide XX example vignettes and code examples on our OSF page/GitHub site for this manuscript across a range of examples of data types provided by the authors of this manuscript. Examples include psycholinguistics, social psychology, COVID related data, and cognitive psychology. The code for this manuscript can be found in *papaja* [@aust2022] formated Rmarkdown document. To
